A Web Application for Wildlife Sightings - How Accessible Is This Technology Using Modern Frameworks?

### 2024-12-18 report/main.tex
Add josisexample.bib for references and update README with project details

% Add 5-10 keywords for every submission
\keywords{
    web application development, 
    wildlife monitoring, 
    geospatial data, 
    user authentication, 
    frontend frameworks, 
    backend programming, 
    database management, 
    rest api integration, 
    mobile web apps, 
    ionic and angular frameworks
    }

@@

The project aimed to evaluate the accessibility of modern technology for creating a web application 
that records geotagged data. Using modern web frameworks, a prototype was developed to document wildlife sightings. 
This application, accessible online and installable on smartphones across major operating systems, 
exemplifies how a straightforward and purpose-driven tool can be built with moderate programming knowledge, 
guided support, and persistence. The project's workflow encompassed backend development in Go, 
frontend design with the Ionic framework, and database integration using MariaDB. Additionally, 
the system incorporates a public data visualization tool and REST API for broader access. 
While the project highlights the potential of current technologies and AI-powered tools like GitHub Copilot, 
it underscores challenges in managing interconnected components. The result is a versatile platform, 
demonstrating both the promise and complexity of custom web application development for geospatial data collection.

@@

This data view is only a frontend to display the data. In order to access the data for further processing, a REST API was created.
A Python script to access the data and a config file for a read-only user is available on the GitHub repository 
for the backend under \path{./db_setup/}.
Feel free to use this script and config file to access the data for testing. However, the data is not valid since it was created for testing purposes only
and anyone testing the app can add random data to the database.

@@

It seems that the technology to create a web application to record geo-tagged data is accessible to everyone with a bit more than basic programming knowledge,
some help, and some insane persistence. The app was successfully developed and deployed. The data view was created and the data can be accessed via a REST API.
Still, it was a real struggle to get to this point, and the hours put into this project are countless. The technology is accessible, but the learning curve is steep.
It is hard to determine if the accessibility is more a result of modern frameworks or the fast-improving AI tools like GitHub Copilot and ChatGPT.

@@

I would like to express my gratitude to Ramon Ott -- the experienced developer -- for his help and support during this project. 
Without his help, this project would not have been possible.
I very much appreciate the help of my fellow student Annalisa Berger for her great work on the design of the logo for the application.

@@

The creation of this document was supported by ChatGPT and GitHub Copilot. 
Additionally, AI played a significant role in the development of the prototype.
The Introduction and the Methods section were translated from German to English using
ChatGPT since they were written before realizing, this report had to follow the JOSIS template,
specifically defining the language. The abstract was initially generated by ChatGPT
and then edited by the author.

>>>>

% Add 5-10 keywords for every submission
\keywords{
    web application development, 
    wildlife monitoring, 
    geospatial data, 
    user authentication, 
    frontend frameworks, 
    backend programming, 
    database management, 
    mobile web apps, 
    ionic and angular frameworks
    }

@@

The project aimed to evaluate the accessibility of modern technology for creating a web application 
that records geotagged data. Using modern web frameworks, a prototype was developed to document wildlife sightings. 
This application, accessible online and installable on smartphones across major operating systems, 
exemplifies how a straightforward and purpose-driven tool can be built with moderate programming knowledge, 
guided support, and persistence. The project's workflow encompassed backend development in Go, 
frontend design with the Ionic framework, and database integration using MariaDB. Additionally, 
the system incorporates a public data visualization tool and direct database access. 
While the project highlights the potential of current technologies and AI-powered tools like GitHub Copilot, 
it underscores challenges in managing interconnected components. The result is a versatile platform, 
demonstrating both the promise and complexity of custom web application development for geospatial data collection.

@@

This data view is only a frontend to display the data. In order to access the data for further processing, a direct database access is available.
A Python script to access the data and a config file for a read-only user is available on the GitHub repository 
for the backend under \path{./db_setup/}.
Feel free to use this script and config file to access the data for testing. However, the data is not valid since it was created for testing purposes only
and anyone testing the app can add random data to the database.

@@

It seems that the technology to create a web application to record geo-tagged data is accessible to everyone with a bit more than basic programming knowledge,
some help, and some insane persistence. The app was successfully developed and deployed. The data view was created and the data can be accessed via a direct database access.
Still, it was a real struggle to get to this point, and the hours put into this project are countless. The technology is accessible, but the learning curve is steep.
It is hard to determine if the accessibility is more a result of modern frameworks or the fast-improving AI tools like GitHub Copilot and ChatGPT.

@@

I would like to express my gratitude to Ramon Ott -- the experienced developer -- for his help and support during this project. 
Without his help, this project would not have been possible.

I very much appreciate the help of my fellow student Annalisa Berger for her great work on the design of the logo for the application.

@@

\begin{itemize}
    \item The creation of this document was supported by ChatGPT and GitHub Copilot.
    \item AI played a significant role in the development of the prototype.
    \item The Introduction and the Methods sections were translated from German to English using ChatGPT 
        since they were written before realizing this report had to follow the JOSIS template, specifically defining the language.
    \item The Abstract was initially generated by ChatGPT and then edited by the author.
    \item A complete spellcheck was performed using ChatGPT.
\end{itemize}
